---
title: "Ch3_Ex_8"
author: "Nibedita Mahalinobis"
format: html
editor: visual
---

This question involves the use of simple linear regression on the Auto data set. Let's have a look at the data set first.

```{r}
#|warning :false
library(ISLR2)
#Viewing a first few rows of dataset
head(Auto)  
#What are all the variables in dataset?
names(Auto)
#What is the nature of the variables?
str(Auto)

```

**Q.8 (a)** Use the lm() function to perform a simple linear regression with mpg as the response and horsepower as the predictor. Use the summary() function to print the results. Comment on the output.

**Solution :**\
We assume the functional form of the regression to be :

$$
           mpg = \beta_0 + \beta_1 \cdot horsepower
$$

```{r}

attach(Auto) #With this command , we specify that all the                 variable names correspond to Auto dataset 

#Fitting the regression 
lm.fit = lm(mpg ~ horsepower)

#To get the summary for the model 
summary(lm.fit) 
```

What can we conclude from the summary given above ?\
**i.** Is there a relationship between the predictor and the response?

**\>** The p value (\< 2.2e-16) for F test is virtually zero . i.e . we can reject the $H_0$ which hypothesizes that there is no relationship present .\
We can conclude that there is some relationship is present between `mpg` and `horsepower` .

**ii.** How strong is the relationship between the predictor and the response?

**\>** As Residual Sum of Errors , RSE is 4.906 , around 21% is the percentage error in prediction of response variable `mpg` . We can see that \$R\^2 = 0.6049\$ which implies that around 60% of variance in `mpg` is explained by the predictor `horsepower.`

```{r}
cor(mpg , horsepower)

```

**iii**.Is the relationship between the predictor and the response
positive or negative?

**\>** Correlation ceofficient is -0.778 indicating that relationship between the predictor and response is negative.

**iv.** What is the predicted mpg associated with a horsepower of
98? What are the associated 95 % confdence and prediction intervals ?

```{r}
#To get confidence interval 
predict(lm.fit , data.frame(horsepower = c(98)) , interval = "confidence")

#To get prediction interval 
predict(lm.fit , data.frame(horsepower = c(98)) , interval = "prediction")
```

**\>** Predicted mpg is 24.26708 with \[ 23.97308 , 24.96108\] as 98% confidence interval and \[14.8094 , 34.12476\] as 98% prediction interval.

**Q.8 (b)** Plot the response and the predictor. Use the `abline()` function to display the least squares regression line.

**Solution :**

```{r}

#To get the scatter plot of variables 
plot(x = horsepower , y = mpg , pch =16 , 
     col = "grey" , main = "Scatterplot of mpg vs horsepower" ) 

#To plot the regression line i.e  line of best fit
abline(lm.fit , col = "blue" , lwd = 2)
```

**Q.8 (c)** Use the plot() function to produce diagnostic plots of the least squares regression ft. Comment on any problems you see with the ft.\
**Solution :**

```{r}
#We first divide the panel into 2 x 2 ,  4 parts 
par(mfrow = c(2,2))

#We plot all the diagnostics plots
plot(lm.fit)
```

**Interpretation** : \
1.Residual Vs Fitted\
Observe that residuals are not equally spread in the Residual vs Fitted plot . We can see some pattern (close to parabolic pattern) , which implies that,there is some problem in linearity assumption. i.e. predictor may have some non linear relationship with response.

2.Normal Q-Q \
If a few observations in the left bottom and right bottom corner are excluded , residuals closely follow the straight line. We can say that residuals are normally distributed.

3.Scale-Location \
Residuals are not equally spread along the ranges of predictors.They are crowed at the right end. Hence, assumption of homoscedasticity(equal variance) is not true.

4.Residuals vs Leverage\
As there points lying beyond the dashed line which is Cook's distance , these are the influential points(leverage points and outliers ) .

To understand more on how to interpret diagnostic plots you can check out the following link: \
<https://library.virginia.edu/data/articles/diagnostic-plots>
